Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)

### 2023-11-14 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

\subsubsection{LUR model evaluation}
Performance of the LUR models was assessed using \textit{k-fold cross-validation}. Specifically, we employed a 10-fold cross-validation technique. In this approach, the data was split into ten distinct sub-samples. For each iteration, the model was trained on k-1 folds and then validated using the leftover fold \cite{ziegel2003}. For every fold, we computed metrics such as \(R^2\), the root-mean-square-error (RMSE), mean bias (MB), and the correlation coefficient (\(r\)). After each iteration, both the error and adjusted \(R\textsuperscript{2}\) were recorded. This process was repeated \(k-times\), and the average of these measures was reported as the indicator of the model validity. 
\vspace{0.5cm}

Details from the metrics used for model validation can be found in the \textbf{Supplementary material X}.

These metrics are defined as follows:
% --- R2 --- %
\begin{equation}
R^2 = \left( \frac{\sum (\hat{y}_i - \bar{y})^2}{\sum (y_i - \bar{y})^2} \right)
\end{equation}

% --- root mean squared error (RMSE) --- % 
\begin{equation}
\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)^2}
\end{equation}

% --- mean bias (MB) --- % 
\begin{equation}
\text{Mean Bias} = \frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)
\end{equation}

% --- correlation coefficient (r) --- %
\begin{equation}
r = \frac{\sum (x_i - \bar{x})(y_i - \bar{y})}{\sqrt{\sum (x_i - \bar{x})^2 \sum (y_i - \bar{y})^2}}
\end{equation}
\vspace{0.5 cm}

@@

The metrics used for the dispersion models we:

% --- Percentage of observed-modelled pairs (n) --- %
\begin{equation}
n = \frac{N_{\text{pairs used}}}{N_{\text{total pairs}}} \times 100
\end{equation}

% --- Factor of Two (FAC2) --- % 
\begin{equation}
\text{FAC2} = \frac{1}{n} \sum_{i=1}^{n} U,
\end{equation}
%Where \( U \) is 1 if \( 0.5 \leq \frac{M_i}{O_i} \leq 2 \) and 0 otherwise.

% --- Mean Bias (MB) --- % 
\begin{equation}
\text{MB} = \frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)
\end{equation}

% --- Mean Gross Error (MGE) --- % 
\begin{equation}
\text{MGE} = \frac{1}{n} \sum_{i=1}^{n} |O_i - M_i|
\end{equation}

% --- Normalized Mean Bias (NMB) --- % 
\begin{equation}
\text{NMB} = \frac{\sum_{i=1}^{n} (M_i - O_i)}{\sum_{i=1}^{n} O_i}
\end{equation}

% --- Root Mean Square Error (RMSE) --- % 
\begin{equation}
\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (O_i - M_i)^2}
\end{equation}

% --- Correlation Coefficient (r) --- % 
\begin{equation}
r = \frac{\sum_{i=1}^{n} (O_i - \bar{O})(M_i - \bar{M})}{\sqrt{\sum_{i=1}^{n} (O_i - \bar{O})^2 \sum_{i=1}^{n} (M_i - \bar{M})^2}}
\end{equation}
\vspace{0.5cm}

>>>>

\subsubsection{LUR model evaluation}
Performance of the LUR models was assessed using \textit{k-fold cross-validation}. Specifically, we employed a 10-fold cross-validation technique. In this approach, the data was split into ten distinct sub-samples. For each iteration, the model was trained on k-1 folds and then validated using the leftover fold \cite{ziegel2003}. For every fold, we computed metrics such as \(R^2\), the root-mean-square-error (RMSE), mean bias (MB), and the correlation coefficient (\(r\)). After each iteration, both the error and adjusted \(R\textsuperscript{2}\) were recorded. This process was repeated \(k-times\), and the average of these measures was reported as the indicator of the model validity. 
\vspace{0.5cm} Details from the metrics used for model validation can be found in the \textbf{Supplementary material X}.

@@
